---
output:
  word_document: default
  html_document: default
---
## ---
## title: "Lascar Calibration and Modelling"
## author: "Prashant Ganesh and Yolanda Hagar"
## date: "March 17 2017"
## ---

Notebook is made more organized with a more organized structure.
Will look into 3 models:
**Model1** : log(Ref) ~ log(lascar) + cummulative + as.factor(lascar number)
**Model2** : Indicator ~ log(reference) + cummulative + as.factor(lascar number)
**Model3** : log(lascar) ~ log(ref) + cummulative + as.factor(lascar number)

For the above 3 models the following techniques will be used:
1) Linear Model
2) Linear Mixed Effects Models (Random Effects Model)
3) General Additive Mixed Model
4) General Additive Model
5) General Linear Model
 
Setting the working directory and loading the file containing the compiled data from MATLAB.

```{r, echo=FALSE}
setwd('~LocalDir...')
library(mixtools)
library(lme4)
require(mgcv)
library(glm2)
library(HI)

d <- read.csv("~/CompiledData.csv") #imports the datafile
d$log.LascarCO = log(d$LascarCO+0.001)
d$log.ReferenceCO = log(d$ReferenceCO + 0.001)
d$CummulativeExposure = d$CummulativeExposure/10000
d$indicator = 0
d$indicator[d$ReferenceCO>0] = 1
LascarOn = as.data.frame(d[which(d$LascarCO>0) ,])
d$sno = 1:nrow(d)



```

# I) First set of models:
## 1 
###Simple Linear Model for *Model1*
```{r}
SimpleModel1 = lm(log.ReferenceCO ~ as.factor(LascarNumber) + log.LascarCO + CummulativeExposure, data = LascarOn) 
summary(SimpleModel1)
```

### General Linear Model for *Model2*
```{r}

GLMModel2 = glm2(indicator ~ CummulativeExposure  + log.LascarCO + as.factor(LascarNumber) , data = d, family = binomial())
summary(GLMModel2)

```

##2)
###Random Effects Models for *Model1*
```{r}
RandomEffectsModel1 = lmer(log.ReferenceCO ~ log.LascarCO+ CummulativeExposure + (1|LascarNumber),data = LascarOn)
summary(RandomEffectsModel1)

```
###Generalized Linear Mixed Models for *Model2*

Using the regular dataset throws the following error: *Some predictor variables are on very different scales: consider rescaling *
http://stackoverflow.com/questions/26904580/error-messages-when-running-glmer-in-r suggests that the datascale be rescaled using lapply()

```{r}

GLMRadnom = glmer(indicator ~ CummulativeExposure2  + log.ReferenceCO + (1|LascarNumber) , data = d, family = binomial())
summary(GLMRadnom)

```

## 3) 
### Generalized Additive Mixed Model for *Model1*
Error saying cannot allocate 32 GB of data
So using bam() General Additive Model for Large Datasets. But dont think it takes Random Effects into consideration
https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/bam.html
```{r}
library(gamm4)
GAMMModel1 = gamm(log.ReferenceCO ~ s(log.LascarCO) + CummulativeExposure, family = gaussian(link=identity), data = LascarOn,random = list(LascarNumber=~1), correlation = corAR1(0.5, form = ~ 1))
summary(GAMMModel1)
#list(LascarNumber=~1)

```

##4) 
### GAMM Model for *Model1* with as.factor(lascar number)

```{r}
GAMModel1 = gam(log.ReferenceCO ~ as.factor(LascarNumber) + s(log.LascarCO) + CummulativeExposure, data = d, family = gaussian(link=identity))

summary(GAMModel1)
```

# II)
## 1
### Linear Model for *Model3*
```{r}

SimpleModel2 = lm(log.LascarCO ~ log.ReferenceCO + CummulativeExposure+ as.factor(LascarNumber)  , data = d) 
summary(SimpleModel2)

```

##2) 
### Random Effects Model for *Model3*
```{r}
RandomEffectsModel3 = lmer(log.LascarCO ~ log.ReferenceCO + CummulativeExposure + (1|LascarNumber),data = d)
summary(RandomEffectsModel3)

```

##3) 
### Generalized Additive *Model3*

```{r}
GAMModel3 = gam( log.LascarCO ~ as.factor(LascarNumber) + s(log.ReferenceCO) + CummulativeExposure, data = d, family = gaussian(link=identity))

summary(GAMModel3)
```

##4) 
### Generalized Additive Mixed Models for *Model3*
```{r}
GAMMModel3 = gamm4(log.ReferenceCO ~ s(log.LascarCO) + CummulativeExposure, family = gaussian(link=identity),random =~ (1|LascarNumber) , data = LascarOn)
summary(GAMMModel3)

```

```{r}

betainit = rep(0,4)
p = 0.5
sigma = 3.275^2
alpha.p = 2
beta.p = 2
n.on = nrow(LascarOn)
n.off = nrow(d) - n.on

nsamples = 100000

#sampledbetas = matrix(0,nrow =as.numeric(nsamples),ncol = 4)
#sampledp =  matrix(0,nrow =nsamples,ncol = 1)
#sampledsigma =  matrix(0,nrow =as.numeric(nsamples),ncol = 1)
sampledbetas = betainit
sampledp = p
sampledsigma = sigma 

c = 10
dd = 10

y = d$LascarCO
x = matrix( c(rep(1,nrow(d)), d$CummulativeExposure , d$ReferenceCO , d$LascarNumber),nrow = nrow(d), ncol = 4) # Have to include lascar Number as factor! 
theta = 0
sigmabeta = 10

tempbeta = betainit


betajposterior = function(x,y,xi,ximinus,betaiminus,sigma,theta,sigmabeta){
  betai = x
 #log.betai.posteriors = - sum((y - xi*betai - ximinus%*%betaiminus)^2) / (2 * sigma^2) - ((betai - theta)^2)/(2*sigmabeta^2)
  betavector = c(betai, betaiminus)
  Xbeta = cbind(xi, ximinus)%*%matrix(betavector, ncol = 1)
 log.betai.posteriors = - sum((y - Xbeta)^2)
  return(log.betai.posteriors)
}


#betaposs = seq(-10, 10, by = .5)
#ii = 2 
#jj = 1
#posterior = NULL
#for(ctr in 1:length(betaposs)){
#posterior = c(posterior, betajposterior(betai = betaposs[ctr], y = y, xi = x[,jj], betaiminus = sampledbetas[ii-1,-c(jj)], ximinus = x[,-c(jj)],sigma = sampledsigma[ii-1],theta = theta, #sigmabeta = sigmabeta))
#}

#summary(posterior)
#plot(betaposs, posterior, type = 'l')


for(ii in 2:nsamples){
  sampledp = c(sampledp,rbeta(1,n.on+alpha.p,n.off + beta.p))
  for(jj in 2:4){
    tempbeta[jj] = arms(sampledbetas[ii-1,jj],betajposterior,function(x) (x>-10)*(x<10), 1, betai = sampledbeta[ii-1,(jj)],y = y, xi = x[,jj] ,ximinus = x[,-c(jj)],
                              betaiminus = sampledbetas[ii-1,-c(jj)],sigma = sampledsigma[ii-1],theta = theta, sigmabeta = sigmabeta)  
  }
  sampledbetas = rbind(sampledbetas,tempbeta)
  sampledsigma = c(sampledsigma, rinvgamm(1,c+n/2,dd+0.5*sum(y - x %*% sampledbeta[ii,])))
}
tempvar = NULL
betaint = c(-10,0,10)
for(ii in 1:3)
{
  
    betas = rep(betaint[ii],4)
    print(- sum((y - x%*%betas)^2))
  
}


sampledbetas
arms(sampledbetas[1,jj],betajposterior, function(x) (x>-10)*(x<10), 1, y = y/10^5, xi = x[,jj]/10^5 ,ximinus = x[,-c(jj)]/10^5,
                              betaiminus = sampledbetas[1,-c(jj)],sigma = sampledsigma[1],theta = theta, sigmabeta = sigmabeta) 
.Call(arms, c(-10, 10), f = function(x) betajposterior(x, y = y, xi = x[,jj] ,ximinus = x[,-c(jj)],
                              betaiminus = sampledbetas[ii-1,-c(jj)],sigma = sampledsigma[ii-1],theta = theta, sigmabeta = sigmabeta), as.integer(1), new.env())





ln = 46
plot(d$LascarTime[which(d$LascarNumber == ln)],d$LascarCO[which(d$LascarNumber == ln)],type = 'l',col = 'red',lwd = 3,xlab = 'Epoch Time in Seconds',
     ylab = 'CO in ppm')
lines(d$LascarTime[which(d$LascarNumber == ln)],d$ReferenceCO[which(d$LascarNumber == ln)],type = 'l',col = 'green',lwd =3)
legend(x = "topright",inset = 0, legend = c('Measured CO' , 'Reference CO'),fill = c('red','green'),cex=1.3,pt.cex = 1)
title(main ='Line Plot of Measured CO vs Reference CO')

```
